---
title: "HW 15"
author: "Nate Lant"
date: "1/23/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r setup, include=F}
library(tidyverse)
library(stats)
library(tseries)
library(forecast)
```

## Problem 2

## Problem 5

## Problem 11 El Nino and Hurricanes. 
Reconsider the El Nino and Hurricane data set from Exercise 10.28.

```{r nino.data, echo=F, message=F}
nino <- read_csv("data/ex1028.csv") %>%
  mutate(Temperature = as_factor(Temperature),
         WestAfrica = as_factor(WestAfrica),
         )
```

Regress the log of the storm index on temperature and the indicator variable for West African wetness
and retain the residuals. 

```{r nino.model, echo=F}
nino.fit <- glm(log(StormIndex) ~ Temperature + WestAfrica, data = nino)
summary(nino.fit)
```

#### (a) Construct a lag plot of the residuals, as in Display 15.5. 
```{r lag.plot, echo=F, warning=F}
# first save the residuals, then create a lag column

# save the residuals
resid.nino <- resid(nino.fit)
#create new columns
nino <- nino %>%
  mutate(residuals = resid.nino,
         resid.lag = lag(residuals))
# pipe into ggplot
nino %>% ggplot(aes(x = resid.lag, y = residuals)) +
  geom_point()
```

Because the lag plot shows even spread, then there is randomness in the data.

#### (b) Construct a partial autocorrelation function plot of the residuals. 

```{r pafc, echo=F}
ggPacf(nino$residuals, 25) +
  ggtitle("Partial Autocorrelation Function Plot")
```

#### (c) Is there any evidence of autocorrelation? How many lags?
No, all of the lags are within the 95% confidence interval.

## Problem 15 Lynx Trappings and Sunspots. 
Many animal populations experience approximately 10-year cycles of boom and bust. An example is the 
Canadian lynx. Annual numbers of lynx trapped in the Mackenzie River district of northwest Canada from 
1821 to 1934 appear in the first five lines of the display below.

```{r display, echo=F, message=F}
lynx <- read_csv("data/ex1515.csv") %>%
  mutate(Year = as_factor(Year))

head(lynx, 5)
```

One conceivable explanation is that these fluctuations are related to solar activity, as described by  
the sunspot series. Is there evidence that lynx numbers are related to sunspot activity in the same 
year or to sunspot activity in the previous one or two years?


## Problem 16



