---
title: "HW 13"
author: "Nate Lant"
date: "1/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```
4 7 15 19 21 22

## 13.4 
Why is there so little interest in how the mean response is associated with blocks in a randomized block experiment?


## 13.7 
If the F-test for treatments is not signiﬁcant but the t-test for one of the contrasts is signiﬁcant, is it proper to report the contrast?

## 13.15
```{r setup.15, include=F}
data_blood <- read_csv("data/case1102.csv") %>% 
  mutate(Time = as_factor(Time),
         ratio = log(Brain/Liver))
```

(a) Ignore the covariates and use an analysis of variance procedure to ﬁt the data. Fit a model that includes interaction terms; plot the residuals versus the ﬁtted values. 

```{r 15.a, echo=F}
glm(ratio ~ Time + Treatment, data = data_blood)
```
 
(b) Test whether there is an interactive effect of treatment and sacriﬁce time. What are the F-statistic, the degrees of freedom, and the p-value? 
 
(c) If there are no interactive effects, test whether there are main effects of treatment and sacriﬁce time. 

(d) Complete the analysis by describing the effects of treatment and sacriﬁce time, either by estimating the appropriate contrasts or by using a regression procedure with indicator variables to model treatment (one indicator) and sacriﬁce time (three indicators). 

## 13.19
## 13.21
hint: class has to be a random effect. 
hint: gain is the response