---
title: "HW 15"
author: "Nate Lant"
date: "1/23/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r setup, include=F}
library(tidyverse)
library(stats)
library(tseries)
library(forecast)
```

## Problem 2

## Problem 5

## Problem 11 El Nino and Hurricanes. 
Reconsider the El Ni˜no and Hurricane data set from Exercise 10.28.

```{r nino.data, echo=F}
nino <- read_csv("data/ex1028.csv") %>%
  mutate(Temperature = as_factor(Temperature),
         WestAfrica = as_factor(WestAfrica),
         )
```

Regress the log of the storm index on temperature and the indicator variable for West African wetness
and retain the residuals. 

```{r nino.model, echo=F}
nino.fit <- glm(log(StormIndex) ~ Temperature + WestAfrica, data = nino)
summary(nino.fit)
```

#### (a) Construct a lag plot of the residuals, as in Display 15.5. 
```{r lag.plot, echo=F}
# first save the residuals, then create a lag column

# save the residuals
resid.nino <- resid(nino.fit)
#create new columns
nino <- nino %>%
  mutate(residuals = resid.nino,
         resid.lag = lag(residuals))
# pipe into ggplot
nino %>% ggplot(aes(x = resid.lag, y = residuals)) +
  geom_point()
```

Because the lag plot shows even spread, then there is randomness in the data.

#### (b) Construct a partial autocorrelation function plot of the residuals. 

```{r pafc, echo=F}
ggPacf(nino$residuals, 25) +
  ggtitle("Partial Autocorrelation Function Plot")
```

#### (c) Is there any evidence of autocorrelation? How many lags?
No, all of the lags are within the 95% confidence interval.

## Problem 15

## Problem 16



