---
title: "HW 15"
author: "Nate Lant"
date: "1/23/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---
```{r setup, include=F}
library(tidyverse)
library(stats)
library(tseries)
library(forecast)
library(lmerTest)
library(pander)
```

## Problem 2 - Can slope predict global warming?
This is a question of serial correlation, which is when a variable is affected by the lagged version of itself.
If serial correlation is present, then the results (including tests and confidence intervals) can be misleading.

## Problem 5 - Breakfast effects
Experiment c) is the best because randomization accounts for a lot of variables that may be lurking.
However, there could be serial correlation because one day's accomplishment should be similar to the previous day.

## Problem 11 - El Nino and Hurricanes. 
Reconsider the El Nino and Hurricane data set from Exercise 10.28.

```{r nino.data, echo=F, message=F}
nino <- read_csv("data/ex1028.csv") %>%
  mutate(Temperature = as_factor(Temperature),
         WestAfrica = as_factor(WestAfrica),
         )
```

Regress the log of the storm index on temperature and the indicator variable for West African wetness
and retain the residuals. 

```{r nino.model, echo=F}
nino.fit <- glm(log(StormIndex) ~ Temperature + WestAfrica, data = nino)
summary(nino.fit)
```

#### (a) Construct a lag plot of the residuals, as in Display 15.5. 
```{r lag.plot, echo=F, warning=F}
# first save the residuals, then create a lag column

# save the residuals
resid.nino <- resid(nino.fit)
#create new columns
nino <- nino %>%
  mutate(residuals = resid.nino,
         resid.lag = lag(residuals))
# pipe into ggplot
nino %>% ggplot(aes(x = resid.lag, y = residuals)) +
  geom_point()
```

Because the lag plot shows even spread, then there is randomness in the data.

#### (b) Construct a partial autocorrelation function plot of the residuals. 

```{r pafc, echo=F}
ggPacf(nino$residuals, 25) +
  ggtitle("Partial Autocorrelation Function Plot")
```

#### (c) Is there any evidence of autocorrelation? How many lags?
No, all of the lags are within the 95% confidence interval.

## Problem 15 Lynx Trappings and Sunspots. 
Many animal populations experience approximately 10-year cycles of boom and bust. An example is the 
Canadian lynx. Annual numbers of lynx trapped in the Mackenzie River district of northwest Canada from 
1821 to 1934 appear in the first five lines of the display below.

```{r display, echo=F, message=F}
lynx <- read_csv("data/ex1515.csv") %>%
  mutate(#Year = as_factor(Year),
         log_Lynx = log(Lynx),
         log_Sunspots = log(Sunspots),
         prev_year = lag(Sunspots),
         prev_2 = lag(prev_year))

lynx %>%
  select(Year, Lynx, Sunspots) %>%
  head(5)
```

One conceivable explanation is that these fluctuations are related to solar activity, as described by  
the sunspot series. Is there evidence that lynx numbers are related to sunspot activity in the same 
year or to sunspot activity in the previous one or two years?

#### Check Assumptions 

Check to see if the distribution in normal. According to the histogram below, a transformation will be necessary. Let's use log(Lynx).

```{r assumptions.lynx, echo=F}
lynx %>% ggplot(aes(x = (Lynx))) +
  geom_histogram()

lynx %>%
  select(Year, log_Lynx, log_Sunspots) %>%
  gather(Type, Value, -Year) %>% 
  #mutate(Year = as.double(Year)) %>%
  # plot
  ggplot(aes(x = Year, y = Value, color = Type)) +
  geom_line()

```

Here is the model. The equation is 
$$ log(Lynx) = \beta_0 + \beta_1*Sunspots + \beta_2*prev_1 + \beta_3*prev_2   $$

```{r spots.years, echo=F}
fit_1 <- glm(log_Lynx ~ log_Sunspots + prev_year + prev_2, data = lynx, na.action=na.exclude)
summary(fit_1)
```

Now from the autocorrelation plot, we see that the last two years are highly significant. This plot shows that the number of lynx is highly correlated with the number of sunspots in the year before and two years before.

```{r pafc.lynx, echo=F}
# store residuals
resid_lynx <- resid(fit_1)

# add residual column to the data set
lynx <- lynx %>%
  mutate(Residuals = resid_lynx)

# ggplot autocorrelation plot
ggPacf(lynx$Residuals) +
  ggtitle("Partial Autocorrelation Function Plot of Lynx sunspots over the years")
```

## Problem 16 - Firearms and Motor vehicles
 Is there evidence of an increasing or decreasing trend in firearm deaths over this
period? What is the rate of increase or decrease? Is there evidence of an increasing or decreasing
trend in motor vehicle deaths over this period? What is the rate of increase or decrease?

```{r data.firearms, echo=F}
# bring in the data from problem 16
data_16 <- read_csv("data/ex1516.csv")

clean <- data_16 %>% 
  gather(Type, Value, -Year) 

#build scatter plot
clean %>%
  ggplot(aes(x = Year, y = Value, color = Type)) +
  geom_line() +
  labs(y = "Deaths (in thousands)") +
  geom_smooth(method = "lm")
```

It looks like there are some trends according to the scatterplot, but let's use a model to see if these slopes are significant.

```{r guns.cars.model, echo=F}
# this is a try at an interaction
# The output is the same as running two regressions. The slope is
# 0.456 (if firearms) and 0.456-0.945=-0.489 (if motordeaths)
glm(Value ~ Year*Type, data = clean) %>%
  summary() %>% pander()


```

Here the output shows that the rate of firearm deaths is increasing by $ 0.456*1000=$ **456** deaths
per year and the rate of motor vehicle deaths is decreasing by $(0.456-0.945)*1000=$ = **489** deaths
per year. These rates are significant on the p<0.001 level.



